The Evolution of Phenotypically Invariant Genetic Networks

### 2017-01-31 kalman_walk.tex
added the abstract and a definition for speciation.

\title{Mathematical Gene Network Evolution}
\maketitle

\begin{abstract}
Abstract.
\end{abstract}

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (0,-4) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (3,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    %\path [->] (A) edge[bend left] node {$c$} (B);
    \path [->] (B) edge node {\tiny $\gamma$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $\lambda_{1}$} (A);
    \path [->] (B) edge[loop left] node {\tiny $\lambda_{2}$} (B);
\end{scope}

\end{tikzpicture}
    \caption{Gene Regulatory Network $A(0)$.}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (0,-4) {$x_{2}$};
    \node[shape=rectangle, dashed] (U) at (3,-2) {$u(t)$};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $(\lambda_{2} - \lambda_{1} - \frac{\gamma}{2})$} (B);
    \path [->, sloped] (B) edge[bend left] node {\tiny $\frac{\gamma}{2}$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $(\lambda_{1} + \frac{\gamma}{2})$} (A);
    \path [->] (B) edge[loop left] node {\tiny $(\lambda_{2}-\frac{\gamma}{2})$} (B);
\end{scope}

>>>>

\title{The Evolution of Phenotypically Invariant Genetic Networks}
\author[1]{Joshua S. Schiffman}
\author[1,2]{Peter L. Ralph}
\affil[1]{University of Southern California, Los Angeles, California}
\affil[2]{University of Oregon, Eugene, Oregon}
\affil[ ]{\texttt{jsschiff@usc.edu} \qquad \texttt{plr@uoregon.edu}}
\maketitle

\begin{abstract}
Bridging the gulf between an organism's genome and phenotype is a poorly understood and complex molecular machinery. Progress in a suite of biological subdisciplines is stalled by our general lack of understanding of this molecular machinery: with respect to both its function and evolution. There does exist a growing body of experiment and data on the evolutionary histories and molecular characterizations of particular gene regulatory networks, as well as thoughtfull verbal and conceptual models. However, as Hardy and Weinberg taught us over a century ago, verbal theories are often insufficient, if not downright misleading. This is especially pertinent given the staggering complexity and scope of contemporary research programmes. This outlook necessitates the advancement of conceptual frameworks of such precision, only mathematics will suffice. Here we outine an analytical theory to study the evolution of biological systems, borrowing insight and methods from control engineering, systems identificaiton, and dynamical systems theory. Presently, we focus on the neutral evolution of genetic regulatory networks. That is, we analytically describe the set of all linear gene networks (of any size) that produce identical phenotypes -- and the evolutionary paths connecting them. In the idealized case of a perfectly adapted population, constant selection, and a static environment, we observe evolution under the ``conservation of phenotype'' as a Brownian motion over phenotypically-invariant network-space. This analysis provides insight to the mechanisms and parameters important for understanding developmental systems drift, network rewiring, evolvability, epistasis, and speciation, as well as the tenuous connection between network architecture and function.

\end{abstract}

@@

\begin{definition}[Biological Species Concept]
  Let $\mathcal{A}$ be the set of all phenotypically symmetric gene networks with transfer funtion $H(z)$. 
  Choose $A$ and $\widetilde A$, both in $\mathcal{A}$. 
  If $\mathcal{H}(A, \widetilde A) \notin \mathcal{A}$ then $A$ and $\widetilde A$ are distinct species. 
\end{definition}

\begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (4,0) {$x_{2}$};
    \node[shape=rectangle] (U) at (2,2) {input};
    \node[shape=rectangle] (y) at (2,-2) {output};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    %\path [->] (A) edge[bend left] node {$c$} (B);
    \path [->] (B) edge node {\tiny $\gamma$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
    \path[->] (A) edge[dashed, bend right] node {\tiny $1$} (y);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $\lambda_{1}$} (A);
    \path [->] (B) edge[loop right] node {\tiny $\lambda_{2}$} (B);
\end{scope}

\end{tikzpicture}
    \caption{Gene Regulatory Network $A(0)$.}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \centering
\begin{tikzpicture}
\begin{scope}[every node/.style={circle,thick,draw}]
  \node (A) at (0,0) {$x_{1}$};
    \node[dashed] (B) at (4,0) {$x_{2}$};
    \node[shape=rectangle] (U) at (2,2) {input};
    \node[shape=rectangle] (y) at (2,-2) {output};
\end{scope}

\begin{scope}[>={Stealth[black]},
              every node/.style={fill=white,circle},
              every edge/.style={draw=black, thick}]
    \path [->, sloped] (A) edge[bend left] node {\tiny $(\lambda_{2} - \lambda_{1} - \frac{\gamma}{2})$} (B);
    \path [->, sloped] (B) edge[bend left] node {\tiny $\frac{\gamma}{2}$} (A); 
    \path[->] (U) edge[dashed] node {\tiny $1$} (A);
    \path[->] (U) edge[dashed] node {\tiny $1$} (B);
    \path[->] (A) edge[dashed, bend right] node {\tiny $1$} (y);
\end{scope}
\begin{scope}[>={Stealth[black]},
              every edge/.style={draw=black, thick}]
    \path [->] (A) edge[loop left] node {\tiny $(\lambda_{1} + \frac{\gamma}{2})$} (A);
    \path [->] (B) edge[loop right] node {\tiny $(\lambda_{2}-\frac{\gamma}{2})$} (B);
\end{scope}
